Actuator line modeling of vertical-axis turbines

### 2016-05-13 paper.tex
Remove paragraphs about momentum and vortex methods

Very fast solution times can be obtained with blade element momentum (BEM)
models, such as the double multiple streamtube (DMST) approached developed by
Paraschivoiu~\cite{Para1988}. However, their poor performance for rotors with
high solidity~\cite{Joo2015} $Nc/R$ (and/or chord-to-radius ratio $c/R$), where
$N$ is the number of blades, $c$ is the chord length, and $R$ is the maximum
rotor radius, makes them less attractive, particular for marine hydrokinetic
applications where rotors of the same size will encounter approximately an order
of magnitude higher torque compared to wind turbines. Their description of the
flow field is very crude as well, simply computing the momentum contained within
discretized streamtubes, which cannot resolve any of the complicated flow
structures shed by the rotor, limiting their applicability for array analysis.

Blade element based vortex line methods, e.g., \cite{Strickland1981,
Murray2011}, can be solved with reasonable computational effort when coupled
with potential flow models---ranging from an expense close to BEM up to that of
2-D blade-resolved CFD. The cost is a function of the number of vortex elements,
which for free vortex methods may increase at each time step, slowing down the
calculation as it marches forward in time, unless there is an additional model
for vortex diffusion. Potential flow vortex methods can resolve more flow
details than BEM. However, since the flow is governed by the Laplace equation,
nonlinear effects such as the turbulent transport will not be included, which we
have determined in experiments to be the same order of magnitude as the mean
vertical advection.

It is therefore desirable to retain a Navier--Stokes description of the flow
field and use an actuator-type model for parameterizing the turbine loading,
which dramatically drives down computational expense by removing the need for
complicated meshes with many fine cells to resolve the boundary layers near the
blade surfaces. As shown in \cite{Bachant2015-JoT}, the conventional uniform
actuator disk is not a good candidate for a cross-flow turbine wake generator,
never mind the fact that it does not typically compute performance predictions.
There are then two actuator methods that use blade element theory to compute
blade loading and therefore generate a wake that more closely resembles that of
an actual turbine. These are the actuator cylinder or swept-surface model (ASSM)
and the actuator line model (ALM). The ASSM solves for the average blade loading
along its path and applies this as a constant body force term in the momentum
equation. The ALM takes a similar approach but is an unsteady method, resolving
the blade element locations in time.

>>>>

For analyzing turbine arrays, it is desirable to retain a Navier--Stokes
description of the flow field---in contrast to, e.g., momentum or potential flow
vortex models---to capture the effects of nonlinear advection and turbulent
transport. However, rather than resolving the fine details in the blade boundary
layers, an actuator-type model for parameterizing the turbine loading may be
employed, which dramatically drives down computational expense. As shown in
\cite{Bachant2015-JoT}, the conventional uniform actuator disk is not a good
candidate for a cross-flow turbine wake generator, never mind the fact that it
does not typically compute performance predictions. There are then two actuator
methods that use blade element theory to compute blade loading and therefore
generate a wake that more closely resembles that of an actual turbine. These are
the actuator cylinder or swept-surface model (ASSM) and the actuator line model
(ALM). The ASSM solves for the average blade loading along its path and applies
this as a constant body force term in the momentum equation. The ALM takes a
similar approach but is an unsteady method, resolving the blade element
locations in time.
